\documentclass[]{article}

%opening
\title{Group Sheet}
\author{}
\date{}
\begin{document}

\maketitle



\section*{Division of labour}

The thematic breakdown of the work is as follows :
\begin{enumerate}
	\item Introduction and electricity market (Jens)
	\item Data Wrangling (Nils)
	\item Benchmark Models (Nils)
	\item Extended Expert Model (Jens)
	\item Seasonal Model (Jens)
	\item Model with external regressors (Nils)
	\item Online learning model (Nils)
	\item Conclusion (Jens) 
	
	
	Online Learning :\\\newline
	Recursive Least Squares is an additional model with OLS. \\\newline
	We estimated the model using the expert model with 2 years \\\newline
	
	Others : \\\newline
	We should think of problems due to an non-full testing/prediction year.\\\newline
	Why did we decide that instead of one year train and the rest as test would not be better than test all\\\newline
	 models with less than a full year. \\\newline
	Why did you even add the MAE errors?\\\newline
	One turbine could roughly produce 2 MW\\\newline
	
	Comments and Questions : \\\newline
	Why is it implausible/plausible to include the renewables in the model?\\\newline
	Connections to the neighbour countries compared \\\newline
	Have a quiet small country with bigger neighbour you should consider those imports of renewables.\\\newline
	We did not include gas because that is information which should not be avaible. \\\newline
	Lagged information for day before the unction so last information of 17:30 \\\newline
	Marginal power plants might set the the price in very heavy loaded hours. \\\newline
	Which other could have be used to improve the models? Why did you use 14 lags ?\\\newline
	Why didn't you preform the Diebold-Mariano-Test for the recursive model? \\\newline
	All estimations are based on least squares this is why we compare RMSE values?\\\newline
	Does outliers have punched into up the MAE/RMSE?\\\newline
	Computational time is negliable.\\\newline
	Lasso could be better because of varibale selection. We did not implement Lasso.\\\newline
	Talking about lags you should add 14 or 15 to check for weekly structures.\\\newline
	We choose the b-spline of 6. What kind of b-spline do you use? How many parameters.\\\newline
	Reduce one due to singularity\\\newline
	What is the b-spline order?\\\newline
	
\end{enumerate}
\end{document}
